Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-03 overleaf/compiler.tex
More cuts.
Finally
Proofread 4.

\figref{fig:compiler_framework} presents an overview of our compiler.
%We start the section with a brief description of the top-level syntax and semantic restrictions in~\secref{sec:syntax}.
%We proceed to describe the front end of the compiler:~\secref{sec:imp_to_ssa}
%describes translation from IMP Source to SSA, and~\secref{sec:ssa_to_MPC_Source}
%describes translation form SSA into MPC Source.
This section outlines the front end, the following section describes analysis on MPC Source and backend vectorization 
and~\secref{sec:backend} outlines the back end. 
%We describe analysis on MPC Source and backend-independent vectorization in~\secref{sec:vectorization}
%and translation into MOTION code in~\secref{sec:backend}. 
Details on all parts appear in~\cite{Anon_TR}.
%\ana{Need to state this more precisely.}

@@

\begin{enumerate}
\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time.
It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
\item Arrays are one-dimensional. $N$-dimensional arrays are linearized and accessed
in row-major order and at this point the programmer is responsible for linearization
and access. (This restriction can be easily lifted.)
\item Array subscrpts are plaintext values as specified by the rule:
\begin{semantics}
    \ntyperule{Array Access}{
%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
  \Gamma \vdash e : \langle \texttt{plain int} \rangle \quad \Gamma \vdash \texttt{A} : \langle q \; \texttt{list}[\tau] \rangle \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
}{
  \Gamma \vdash \texttt{A[$e$]} : \langle q \; \tau \rangle
}
\end{semantics}
The subscript $e$ is a function of the indices of the enclosing loops.
For read access, the compiler allows an arbitrary such function.
However, it restricts write access to \emph{canonical writes}, i.e., {\sf A[$i,j,k$] = ...}
where $i$, $j$ and $k$ loop over the three dimensions of {\sf A}.
Write access such as for example {\sf A[$i,j+2$] = ...} is not allowed. (This is again a restriction we imposed for convenience in our current implementation; we plan to extend the compiler with arbitrary write access.)
%\item \ana{Add rules for logical ops}
%\item \ana{Add rules for arithmetic ops}

>>>>

\figref{fig:compiler_framework} presents an overview of our compiler.
%We start the section with a brief description of the top-level syntax and semantic restrictions in~\secref{sec:syntax}.
%We proceed to describe the front end of the compiler:~\secref{sec:imp_to_ssa}
%describes translation from IMP Source to SSA, and~\secref{sec:ssa_to_MPC_Source}
%describes translation form SSA into MPC Source.
This section outlines the front end, \secref{sec:vectorization} describes analysis on MPC Source and backend-independent vectorization 
and~\secref{sec:backend} outlines the back end. 
%We describe analysis on MPC Source and backend-independent vectorization in~\secref{sec:vectorization}
%and translation into MOTION code in~\secref{sec:backend}. 
Details on all parts appear in~\cite{Anon_TR}.
%\ana{Need to state this more precisely.}

@@

\begin{enumerate}
\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time.
It is a standard restriction in MPC that computation is bounded~\cite{NDSS:DemSchZoh15,CCS:BDKKS18}.
\item Arrays are one-dimensional. $N$-dimensional arrays are linearized and accessed
in row-major order and at this point the programmer is responsible for linearization
and access. (This restriction can be easily lifted.)
\item Array subscrpts are plaintext values as specified by the rule:
\begin{semantics}
    \ntyperule{Array Access}{
%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
  \Gamma \vdash e : \langle \texttt{plain int} \rangle \quad \Gamma \vdash \texttt{A} : \langle q \; \texttt{list}[\tau] \rangle \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
}{
  \Gamma \vdash \texttt{A[$e$]} : \langle q \; \tau \rangle
}
\end{semantics}
The subscript $e$ is a function of the indices of the enclosing loops.
For read access, the compiler allows an arbitrary such function.
However, it restricts write access to \emph{canonical writes}, i.e., {\sf A[$i,j,k$] = ...}
where $i$, $j$ and $k$ loop over the three dimensions of {\sf A}.
Write access such as for example {\sf A[$i,j+2$] = ...} is not allowed. (This is again a restriction we imposed for convenience in our current implementation; we plan to extend the compiler with arbitrary write access.)
%\item \ana{Add rules for logical ops}
%\item \ana{Add rules for arithmetic ops}

### 2022-05-03 overleaf/vectorization.tex
Proofread 4.

This section describes our vectorization algorithm. While vectorization is a longstanding problem, 
and we build upon existing work on scalar expansion and classical loop vectorization~\cite{Allen:1987}, 
our algorithm is unique as it works on the MPC Source SSA-form representation. We posit that vectorization 
over MPC Source is a new problem that warrants a new look, in part because of MPC's unique linear 
structure and in part because vectorization meshes in with other MPC-specific optimizations in non-trivial ways
(other works have explored manual vectorization and protocol mixing in an ad-hoc way, e.g.,~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Ishaq:2019}).

In this section, we briefly describe the analysis. Details, including edge cases and examples, can be found in the supplementary extended version~\cite{Anon_TR}.
%\secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,
%which lifts scalars (and arrays) to the corresponding loop dimensionality to create opportunities for vectorization. 
%\secref{sec:basic_vectorization} describes our core vectorization algorithm and~\secref{sec:correctness} argues correctness of the vectorization transformation.
%\secref{sec:array_writes} extends vectorization with array writes.

>>>>

This section describes our vectorization algorithm. While vectorization is a longstanding problem, 
and we build upon existing work on scalar expansion and classical loop vectorization~\cite{Allen:1987}, 
our algorithm is unique as it works on the MPC Source SSA-form representation. We posit that vectorization 
over MPC Source is a new problem that warrants a new look, in part because of MPC's unique linear 
structure and in part because vectorization meshes in with other MPC-specific optimizations in non-trivial ways
(other works have explored manual vectorization and protocol mixing in ad-hoc ways, e.g.,~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Ishaq:2019}).

In this section, we briefly describe the analysis. Details, including edge cases and examples, can be found in the extended version~\cite{Anon_TR}.
%\secref{sec:dependence} describes our dependence analysis and ~\secref{sec:expansion} describes scalar expansion,
%which lifts scalars (and arrays) to the corresponding loop dimensionality to create opportunities for vectorization. 
%\secref{sec:basic_vectorization} describes our core vectorization algorithm and~\secref{sec:correctness} argues correctness of the vectorization transformation.
%\secref{sec:array_writes} extends vectorization with array writes.
